---
title: "Aides Drug"
author: "Teresa Juan"
date: "06/14/22"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```


```{r eval=FALSE}
triUmp <- triUmp %>%
mutate(across(where(is.character), as.factor))
```

```{r}
females <-filter(triUmp, gender == "F")
```



# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

In the data we can see that 60% of the females that took the drug for aids end up getting it then the ones that took the placebo. However not allot of females  got aids from the 1,000 trials just a few of them, but the one that took the drug for aids shows us that it didn't work. It seems that the drug made them more active, and the placebo had no effect.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```

As we look at the table the results showed us that females that took the treatment with drug had a 48% percent catching aids, and the ones that took the placebo cough aids 41%. We can see that the ones that took the placebo had a really low chance of getting aids, then those that took the drug.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chances of this happening  based on the P value its a 7 out of 10,000 chances. If the drug worked that would have been a positive result, however this test did the opposite, more females were infected with aids as they took the drug then those that took the placebo.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In conclusion as we worked trough the data and the fisherXY test, we immediately notice that this clinical trial on females catching aids if they took the drug was a fail. We notice that females  that took the drug did in fact are more at risk to catch the disease than those that took the placebo. As we looked at the table we see that almost 60% of the 100% that took the drug got infected with aids, and only 40% of them didn't. The fisherXY test showed us that you are more likely to catch aid if you take the drug then if you take the placebo. Based on the p value  of .00006596 being it such a small number proved that the drug wasn't effective against AIDS. That's a chance that now one would like to take based on this test, the clinical trial was a fail.







